An Overview on Temporal Databases.\\ New Approaches and Proposals.

### 2012-02-27 trunk/book/sources/frw.tex
Last draft of the paper. This is the final draft.

In this work we have presented the main properties of the time in an information system. An overview of the commercial temporal DBMS has also been introduced. Although, to the best of our knowledge, no commercial fuzzy temporal DBMS have been found.  Finally the managing of imprecision in temporal databases is explained. A brief introduction of bipolar querying in temporal databases is done at the end of the section.

Further research work should be done in several ways: first of all, a theoretical model for both dealing with uncertainty in the database as well as in the querying should be defined. Then, another interesting work is to propose an implementation with both DDL (Data Definition Language) and DML (Data Manipulation Language).

>>>>

In this chapter, some of the main concepts concerning temporal modelling and temporal modelling in information systems and the terminology corresponding with these concepts are introduced and explained and some of the main properties of and issues with these concepts are presented and discussed. An overview of some commercial temporal DBMS is briefly introduced. Finally, a novel technique for querying valid-time relations using imperfect query specifications is presented.

 %we have presented the main properties of the time in an information system. An overview of the commercial temporal DBMS has also been introduced. Although, to the best of our knowledge, no commercial fuzzy temporal DBMS have been found.  Finally the managing of imprecision in temporal databases is explained. A brief introduction of bipolar querying in temporal databases is done at the end of the section.

Further research work could follow several general directions. First of all, a theoretical model for dealing with uncertainty in both the database and the query at the same time could be researched and defined. Next, implementations including both DDL and DML could be proposed and constructed.

%Further research work should be done in several ways: first of all, a theoretical model for both dealing with uncertainty in the database as well as in the querying should be defined. Then, another interesting work is to propose an implementation with both DDL (Data Definition Language) and DML (Data Manipulation Language).

### 2012-02-27 trunk/book/sources/fuzzyTemporalDatabases.tex
update with necessity in allen relations
Last draft of the paper. This is the final draft.

In the presented approach, for every record $r$, each part of $\tilde{Q}$ is evaluated independently:
\begin{itemize}
\item
The user preferences expressed in the non-temporal part, $Q$, are evaluated, resulting in a satisfaction degree denoted as $e_Q(r)$. The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
\item
The evaluation of the temporal demand expressed in the temporal part, $Q^{time}$, depends on $AR$. A specific set of ill-known constraints is considered depending on the Allen relation denoted by $AR$. The exact set chosen can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$. Then both possibility and necessity measures are computed and aggregated.
\end{itemize}

@@

\centering
\begin{tabular}{|c|c|}
\hline
Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
\hline
I before J & $\sup_{a>w}\pi_x(w)$\\
\hline

\multirow{2}{*}
{I equal J} &  $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
 &  $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
\hline

I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
\hline

\multirow{2}{*}
{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ \\
 & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
\hline

\multirow{2}{*}
{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
 & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
\hline

{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
\hline

{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
\hline

@@

\subsubsection{Ranking and aggregation}
For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained. Then, the measures are combined in order to obtain an evaluation score $e_{Q_{time} \left( r \right)}$. The evaluation function is computed by:

\begin{equation}
\label{eq:temporal-eval}
e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
\end{equation}

This measure provides a natural score because of the following property of the necessity measure: 
\begin{equation}
\label{eq:necessity} %iff or iif?
\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
\end{equation}

The final ranking is given by a convex combination of both temporal and non-temporal evaluations:

\begin{equation}
\label{eq:convex-combination}
e_{final} \left( r \right) = \left \lbrace t \mid t = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}, t,\omega \in \left[0,1\right] \right \rbrace
\end{equation}

\subsection{Bipolarity in temporal databases}
\label{subsubsec:bipolarity}
Humans express their preferences in both positive and negative statements. This property is interesting in querying of databases, because sometimes, the user does not know exactly his / her preferences but what he / she does not want. \emph{Bipolarity} is therefore a technique to model both positive and negative preferences. Sometimes positive and negative information are clearly symmetric and therefore, it is possible to derive one information from another e.g. a person may define the concept of tall as 1.80 meters or higher. The negative would be the opposite, not tall is therefore, being less than 1.80 meters.  But in some cases, positive information can not be obtained directly from the negative information and vice-verse. E.g. a person wants to buy a blue motorbike, this does not mean that the person would reject a black motorbike. This heterogeneous semantics are given by the \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.\\

Some previous work have been done in ~\cite{Lacroix87}. There, instead of positive and negative criteria, desired and mandatory query conditions are explained. This can be seen as a way to specify positive and negative information: the inverse of a mandatory condition is what should be rejected (negative information), whereas desired conditions can be seen as positive information.\\

This introduces the necessity of a bipolar way of querying databases. Therefore, the bipolar querying of databases~\cite{DeTre2009} is a querying technique which allows the user to express both positive and negative criteria in the database query.\\

Bipolarity is handled by means of different tools:

\begin{itemize}
\item
The intuitionistic fuzzy sets  \emph{AFS} by Atannasov ~\cite{Atanassov1986}. The approach re-define the concept of inverse. Instead of considering the inverse of a set just the complementary, a AFS is defined by means of a membership function and a non-membership function.
\item
The interval valued fuzzy sets \emph{IVFS} were introduced independently by Zadeh~\cite{Zadeh75a}, Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75} in the seventies~\cite{Dubois05}. %look for the reference!
\item
The two fold fuzzy sets \emph{(TFS)} by Dubois and Prade~\cite{Dubois02}. This approach defines desired and mandatory query conditions.
\end{itemize}

From a theoretical point of view, in a database, bipolarity may be found at:
\begin{description}
\item[\textbf{The query}] It is possible to distinguish between:
	\begin{itemize}
	\item
	Inside criterion bipolarity: Each individual criteria may be specified with both positive and negative conditions. E.g. the user wants a black car, but definitively not a red neither a blue one.
	\item
	Outside criterion bipolarity: The query is specified with a global positive and negative conditions. E.g. The user wants a black car with a fuel consumption of 5 liters or less but definitively not a white car with a fuel consumption of 6 liters or more. 
	\end{itemize}
\item[\textbf{The data storage}] In the database some attributes or even at tuple level it should be possible to specify them with bipolarity. Nevertheless, there is not so many research on this topic.
%	\begin{itemize}
%	\item
%	Attribute bipolarity: Each attribute may be specified  
%	\item
%	Tuple bipolarity:
%	\end{itemize}
\end{description}

A proposal for the bipolar querying of temporal databases has been made by Billiet~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. First of all, the query structure is shown. Then, both evaluation and ranking methods are defined. The model deals with a fuzzy valid-time specification in~\cite{garrido2009}.

>>>>

In the presented approach, for every record $r$, each part of $\tilde{Q}$ is evaluated independently:
\begin{itemize}
\item
The user preferences expressed in the non-temporal part, $Q$, are evaluated, resulting in a satisfaction degree denoted as $e_Q(r)$. The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
\item
The evaluation of the temporal demand expressed in the temporal part, $Q^{time}$, depends on $AR$. A specific set of ill-known constraints\cite{Pon11} is considered depending on the Allen relation denoted by $AR$. The exact set chosen can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$. Then, using these formulas given in table \ref{tab:fuzzy-allen-relations}, both the possibility degree $\Pos_{Q^{time}}(r)$ and the necessity degree $\Nec_{Q^{time}}(r)$ are computed.
\end{itemize}

\textcolor{red}{TODO: maybe part on what these and mean? Maybe part on where the formulas in table \ref{tab:fuzzy-allen-relations} come from?}

@@

\centering
\begin{tabular}{|c|c|c|}
\hline
Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ & $\Nec\left(C_{i}, i = 1,2,... \right)$ \\
\hline
I before J & $\sup_{a>w}\pi_X(w)$ & $\inf_{a \leq w} 1- \pi_X(w)$\\
\hline

\multirow{4}{*}
{I equal J} &  $\min ( \sup_{a \leq w}\pi_X(w),$  & $\min ( \inf_{a>w} 1-\pi_X(w),$\\
& $\pi_X(w),$ & $\inf_{w \in \left[a,b\right]} \pi_X(w), $ \\
 &  $\sup_{b \geq w}\pi_Y(w),$ & $\inf_{b>w} 1-\pi_Y(w),$\\
 & $\pi_Y(w))$ & $\inf_{w \in \left[a,b\right]} \pi_Y(w) )$ \\
\hline

\multirow{2}{*}
I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ & $\min (\inf_{a<w} 1 - \pi_X(w),$\\
& $\pi_X(w))$  & $ \inf_{w \in \left[a,b\right]} \pi_X(w) )$ \\
\hline

\multirow{3}{*}
{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ & $\min ( \inf_{b \geq w} 1 - \pi_Y(w),$\\
 & $\sup_{a \geq w}\pi_X(w),$& $1- \inf_{a>w}1-\pi_X(w),$\\
 & $\sup_{a \leq w}\pi_X(w))$ & $1- \inf_{a<w}1-\pi_X(w) )$ \\
\hline

\multirow{4}{*}
{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),,$ & $\max (\min (\inf_{a \leq w}1-\pi_X(w), $\\
 & $\sup_{b \geq w}\pi_Y(w)) $ & $ \inf_{b>w} 1- \pi_Y(w)),$ \\
 & $\min ( \sup_{a \leq w }\pi_X(w),$ & $\min ( \inf_{a<w}1-\pi_X(w),$\\
 & $\sup_{b>w}\pi_Y(w))$ & $\inf_{b\geq w}1-\pi_Y(w)))$ \\
\hline
\multirow{2}{*}
{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$  & $\min (\inf_{a<w}1-\pi_X(w),$\\
& $\pi_X(w))$ & $inf_{w \in \left[a,b\right]} 1- \pi_X(w))$ \\
\hline
\multirow{2}{*}
{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$  & $\min(\inf_{b>w}1-\pi_Y(w),$\\
& $\pi_Y(w))$ & $\inf_{w \in [a,b]} 1- \pi_Y(w))$ \\
\hline

@@

\subsubsection{Aggregation and Ranking}
To be able to present the most appropriate results to the user, for every record $r$, an aggregation method is used to aggregate $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ into a temporal record rank $e_{Q^{time}}(r)$ and after this, a convex combination combining $e_{Q^{time}}(r)$ and $e_{Q}(r)$ will provide the final record rank $e_{final}(r)$.

%For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained. Then, the measures are combined in order to obtain an evaluation score $e_{Q_{time} \left( r \right)}$. The evaluation function is computed by:

To calculate $e_{Q^{time}}(r)$, an a simple and crude method is used:

\begin{equation}
\label{eq:temporal-eval}
e_{Q^{time}}(r) = \left(\frac{\Pos_{Q^{time}}(r)+\Nec_{Q^{time}}(r)}{2}\right)
%e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
\end{equation}

This method aims to provide the result records with a natural ranking based on the users temporal constraint. $e_{Q^{time}}(r)$ will of course be a value in $\left[0,1\right]$, as both $\Pos_{Q^{time}}(r) \in \left[0,1\right]$ and $\Nec_{Q^{time}}(r) \in \left[0,1\right]$. The purpose is that records which fit the users temporal demand better get a higher score than records fitting the temporal demand worse. Here, this aim is reached because the necessity degree $\Nec_{Q^{time}}(r)$ cannot exceed $0$ unless the possibility degree $\Pos_{Q^{time}}(r)$ equals $1$.

%This measure provides a natural score because of the following property of the necessity measure: 
%\begin{equation}
%\label{eq:necessity} %iff or iif?
%\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
%\end{equation}

The final ranking $e_{final}(r)$ for a record $r$ is now given by a convex combination of both temporal and non-temporal evaluation scores.

\begin{equation}
\label{eq:convex-combination}
e_{final} \left( r \right) = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}
\end{equation}

A convex combination is used mainly for 2 reasons:
\begin{itemize}
	\item The use of this convex combination allows a record to make up for a low temporal evaluation score with a high non-temporal evaluation score and vice versa.
	\item The exact value of $\omega$ can now be modified to ascribe more value to either the fulfillment of the user's temporal demands or the fulfillment of the user's non-temporal constraints.
\end{itemize}

\subsection{Bipolarity in temporal databases}
\label{subsubsec:bipolarity}
Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable \cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. This realization is interesting with regard to database querying, because sometimes a user does not exactly know his or her preferences or can't express them in only positive statements, but prefers to use negative statements to express what he or she dislikes or doesn't need. This introduces the need for \emph{bipolar querying}, a technique to model both positive and negative user preferences in a database query. Sometimes positive and negative preferences are clearly symmetric, making it possible to derive one from the other. For example, a person may define the concept of `tall' as `1.80 meters or higher'. The negative would then be the opposite: not tall would be `anyhing less than 1.80 meters'. However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa. E.g. when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike. This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.

The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature \cite{DeTre2009}, \cite{Dubois2008}, \cite{Lacroix87}. In \cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences. However, the inverse of a mandatory preference expresses what should be rejected and this could be seen as negative information, whereas desired query conditions can be seen as positive preferences. However, the combination of bipolar querying and the use of imprecise query preferences in the context of temporal databases is not so well discussed in existing literature. A proposal for the bipolar querying of valid-time databases has been made by Billiet et al.~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. The model presented there deals with a fuzzy valid-time specification based on~\cite{garrido2009}.

%Some previous work have been done in ~\cite{Lacroix87}. There, instead of positive and negative criteria, desired and mandatory query conditions are explained. This can be seen as a way to specify positive and negative information: the inverse of a mandatory condition is what should be rejected (negative information), whereas desired conditions can be seen as positive information.\\

%This introduces the necessity of a bipolar way of querying databases. Therefore, the bipolar querying of databases~\cite{DeTre2009} is a querying technique which allows the user to express both positive and negative criteria in the database query.\\

Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets \cite{Dubois02}.

%\begin{itemize}
%\item
%Intuitionistic Fuzzy Sets~\cite{Atanassov1986} are sometimes used. The approach re-define the concept of inverse. Instead of considering the inverse of a set just the complementary, a AFS is defined by means of a membership function and a non-membership function.
%\item
%The interval valued fuzzy sets \emph{IVFS} were introduced independently by Zadeh~\cite{Zadeh75a}, Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75} in the seventies~\cite{Dubois05}. %look for the reference!
%\item
%The two fold fuzzy sets \emph{(TFS)} by Dubois and Prade~\cite{Dubois02}. This approach defines desired and mandatory query conditions.
%\end{itemize}

From a theoretical point of view, bipolarity might be found either in the queries presented to a database system or in a database managed by a database system.

When bipolarity is found in queries, it is possible to distinguish between:
\begin{itemize}
	\item
	Bipolarity inside query criteria: each individual query criterion may be specified using both positive and negative preferences. For example when querying a car database, the user can express that he or she wants a black car, but definitely not a red neither a blue one. Bipolarity resides here within the car color criterion.
	\item
	Bipolarity outside query criteria: the query is specified using a global positive and a global negative preference part. For example when querying a car database, the user can express that he or she wants a black car, but definitely not a car with a fuel consumption of 6 liters or more.
	\end{itemize}

Concerning bipolarity inside a database, it should be possible to specify both positive and negative real world object or concept aspects, even at record level. Nevertheless, not so much research exists concerning bipolarity in databases.

%some attributes or even at tuple level it should be possible to specify them with bipolarity. Nevertheless, there is not so many research on this topic.
%	\begin{itemize}
%	\item
%	Attribute bipolarity: Each attribute may be specified  
%	\item
%	Tuple bipolarity:
%	\end{itemize}

### 2012-02-27 trunk/book/sources/temporalDatabases.tex
Last draft of the paper. This is the final draft.

@@

%\vspace{-25pt}
This latter issue is the concern of the next subsection.

\subsection{\label{subsubsec:consistency}Consistency in Temporal Relations}
As stated in subsection \ref{subsubsec:primary-key}, one of the main issues concerns the insertion of spurious values in a relation. The most usual approach to deal with this is to change the database DML to enforce the consistency with which the temporal relation models the temporal properties of the part of reality the relation models. An example of such DML adaptation is given below.

@@

\begin{example}
\textcolor{red}{PROBLEM}
Consider again the example relation visualized in table \ref{table:example-database}. Imagine John left the company, but nobody updated the relation, so it is not known when

@@

\begin{verbatim}
Update Employees set StartYear = 2010 where ID=3;
\end{verbatim}

Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version. The update sentence showed above is translated into:

\begin{verbatim}
Update Employees set EndYear = 2010 where ID=3;
Insert into Employees values (3,John,52,-,2010,-);
\end{verbatim}

@@

\begin{table}
\centering
\caption{Example relation from example \ref{ex:pk} updated maintaining consistency.}
\begin{tabular}{c c c c c c }
\hline
\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
1 & Peter & 24 & John &  2010 & - \\
2 & Maria & 34 & John & 2001 & - \\
3 & John & 52 & - &  1999 & 2010 \\
3 & John & 52 & - &  2010 & - \\
4 & Sarah & 29 & Maria &  2005 & 2009 \\
\hline 
\end{tabular}
\label{table:example-database-update}

>>>>

\begin{note}
In the following tables (\ref{table:example-database},\ref{table:example-database-with-new-pk},\ref{table:example-database-update}) the relation `Supervisor' stores the ID of the employee. To improve the readability of the example, the tables show the name of the employee. Note also that the field for the age is expressed as the age in years whereas the value stored in the database is the birthday.
\end{note}

@@

%\vspace{-25pt}
The next subsection is concerned with another interresting issue.

\subsection{\label{subsubsec:consistency}Consistency in Temporal Relations}
As stated in subsection \ref{subsubsec:primary-key}, one of the main issues concerns the insertion of spurious values in a relation. Another issue concerns the possible breach of consistency when updating records. The most usual approach to deal with this is to change the database DML to enforce the consistency with which the temporal relation models the temporal properties of the part of reality the relation models. An example of such DML adaptation is given below.

@@

\begin{example}
Consider the database in table \ref{table:example-database}. The employee with ID=3 (John) works now for the employee with ID= 4 (Sarah). The update is made in the following sentence:

@@

\begin{verbatim}
Update Employees set 'Works for' = 4 where ID=3;
\end{verbatim}

Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version. Consider that the time in the  system is the year 2010. Then, the update sentence showed above is translated into:

\begin{verbatim}
Update Employees set EndYear = 2010 where ID=3;
Insert into Employees values (3,John,52,Sarah,2010,-);
\end{verbatim}

@@

\begin{table}
\centering
\caption{Example relation from example \ref{ex:pk} updated maintaining consistency.}
\begin{tabular}{c c c c c c }
\hline
\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
1 & Peter & 24 & John &  2010 & - \\
2 & Maria & 34 & John & 2001 & - \\
3 & John & 52 & - &  1999 & 2010 \\
3 & John & 52 & Sarah &  2010 & - \\
4 & Sarah & 29 & Maria &  2005 & 2009 \\
\hline 
\end{tabular}
\label{table:example-database-update}

### 2012-02-27 trunk/book/sources/timeDomain.tex
Last draft of the paper. This is the final draft.

\textcolor{red}{MOVE TO CHAPTER 4}
The transition between granularities~\cite{Lin97} is also considered as a source of imprecision. In a sentence like \emph{`The Alhambra is 612 years old'}, the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity. However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed. Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.

\textcolor{red}{END MOVE TO CHAPTER 4}

@@

Several operators are defined in order to compare temporal elements or time points. Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points. Figure \ref{fig:allen} shows the temporal relationships Allen discerned. Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.

\textcolor{red}{MOVE TO FUZZY PART}
For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easily implemented by means of regular fuzzy operations.
\textcolor{red}{END MOVE TO FUZZY PART}

>>>>

%\textcolor{red}{MOVE TO CHAPTER 4}
%The transition between granularities~\cite{Lin97} is also considered as a source of imprecision. In a sentence like \emph{`The Alhambra is 612 years old'}, the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity. However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed. Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.

%\textcolor{red}{END MOVE TO CHAPTER 4}

@@

Several operators are defined in order to compare temporal elements or time points. Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points. Figure \ref{fig:allen} shows the temporal relationships Allen discerned. Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
